### 2022-02-03 Manuscript/StandardizedEvaluation.tex
Add kable table

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Towards a standardized evaluation of multiple imputation routines},
  pdfauthor={Hanne Oberman and Gerko Vink},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Towards a standardized evaluation of multiple imputation
routines}
\author{Hanne Oberman and Gerko Vink}
\date{19-1-2022}

\begin{document}
\maketitle
\begin{abstract}
Developing new imputation methodology has become a very active field.
Unfortunately, there is no consensus on how to perform simulation
studies to evaluate the properties of imputation methods. In this paper
I propose a move towards a standardized evaluation of imputation
methods. To demonstrate the need for standardization, I highlight a set
of potential pitfalls that bring forth a chain of potential problems in
the objective assessment of the performance of imputation routines. This
may lead to suboptimal use of multiple imputation in practice.
Additionally, I suggest a course of action for simulating and evaluating
missing data problems.
\end{abstract}

\hypertarget{to-do}{%
\section{TO DO}\label{to-do}}

\begin{itemize}
\tightlist
\item
  add ADEMP paper (Morris, White, and Crowther 2019), diagnostic
  checking paper (Zhao 2022)
\item
  add PPC (refer to Mingyang, and Gelman \& King)
\item
  expand convergence section
\item
  add RMSE
\item
  add insights from literature review for chapter on missing data in EHR
\item
  add link to simulation script, and \(n\)
\item
  add simulation results
\end{itemize}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Multiple imputation Donald B. Rubin (1987) is a state-of-the-art
technique for drawing valid conclusions from incomplete data, and is
increasingly used in many substantive fields, such as medicine (see e.g.
Pulkki-Råback et al. 2015; Alkire et al. 2015; Stinesen Kollberg et al.
2015), psychology (e.g. Sorbi et al. 2014), economics (e.g. Wong et al.
2015; Dafoe, Oneal, and Russett 2013; Allan and Forder 2015), sociology
(e.g. Crowley and Kolenikov 2014; De Hoon and Van Tubergen 2014), and
official statistics (e.g. Roderick J. Little 2012; Klausch, Hox, and
Schouten 2015).

The technique has earned a permanent spot in research and policy making,
demonstrated e.g.~by the detailed manual created by the National
Research Council (Roderick J. Little et al. 2012). Although a top-down
enforcement of valid ways to handle missing data is not yet very
pronounced, an increasing amount of researchers are embracing multiple
imputation techniques. After all, the principle of multiple imputation
is very intuitive.

The idea behind multiple imputation is to impute (fill in) the missing
values multiple times to obtain a valid estimate of what could have
been. The multiple datasets that are thus obtained can be analyzed by
standard techniques and the analysis results can be combined into a
single inference. In contrast to ad hoc methods for dealing with missing
values (e.g.~list-wise deletion, mean imputation, regression imputation,
last observation carried forward), multiple imputation properly takes
the sources of uncertainty that are related to the missingness problem
into account.

The quality of a solution obtained by multiple imputation depends on the
statistical properties of the incomplete data and the degree to which an
imputation procedure is able to capture these properties when modeling
missing values. In general it holds that modeling missing data becomes
more challenging when the amount of missingness increases. However, when
(strong) relations in the data are present, the observed parts can hold
great predictive power for the models that estimate the missingness. In
that case, multiple imputation would be substantially more efficient
than the ubiquitous list-wise deletion.

When evaluating the statistical properties (and thereby the practical
applicability) of new imputation methodology, researchers most often
make use of simulation studies. In such studies a complete data set is
usually generated from a statistical model, another model is used to
induce missingness, and a set of evaluation criteria is postulated to
evaluate the performance of the imputation method. However, no golden
standard has been established and, as a result, validity of the
simulation and evaluation procedures may differ tremendously from one
developer to another.

The purpose of this paper is threefold: First, I raise some concerns
with respect to evaluating imputation methodology. These concerns stem
from careful consideration with fellow `imputers' and from encounters as
a reviewer for statistical journals. Second, I provide imputation
methodologists with a suggested course of action when simulating missing
data problems. This suggested approach should identify a common ground,
but is in no way intended as an absolute solution. This identifies the
third purpose of this paper: discussion. I hope to elicit critical
thinking with respect to the problems at hand. We are all convinced that
our methodology has some merit. But for sake of progress it would be
much more advantageous if the aim of our evaluations would go beyond
\emph{proving the point} and would legitimately consider the statistical
properties.

\hypertarget{why-some-evaluations-should-not-be-trusted}{%
\subsection{Why some evaluations should not be
trusted}\label{why-some-evaluations-should-not-be-trusted}}

As of today, there is no consensus on how to perform simulation studies
to evaluate the properties of imputation methods. Typically, the
developer of a new imputation routine does some tests by simulations,
but these tests differ across developers. This brings forth a chain of
potential problems in the objective assessment of the performance of
imputation routines that may lead to suboptimal use of multiple
imputation in practice. To demonstrate the broad impact of these
problems, I subdivide the problems in the following three distinct
categories: problems with data generation, problems with missingness
generation and problems with performance evaluation. I further detail
the impact these problems may have on the validity of the performance
evaluation of the imputation routines.

\hypertarget{data-generation-problems}{%
\subsection{Data generation problems}\label{data-generation-problems}}

To evaluate the ability of an imputation routine to handle the
non-response model, a form of truth has to be established. Those who
perform simulation studies are in the luxury position to establish the
truth beforehand, either by generating data from a statistical model, or
by sampling data from a sufficiently large observed set. However, when
generating data, two easily overlooked problems may arise.

First, data are often generated following the model (or distribution)
that is also used for imputing the data. The performance of the
imputation approach is then deemed good, which is no surprise as the
evaluated conditions are in favor of the problem that is studied.
Although such procedures may be statistically relevant, the approach
would be no good for even the simplest case. Such simulations pose
serious threats to the applicability of the method as real-life data
hardly ever follow a given theoretical distribution.

Second, data are often generated such that the problem that is being
studied is most pronounced. This results in simulated data that contains
very valuable information structures, i.e.~the correlations between
groups of variables may be very pronounced. In other words, no matter
what type of missingness is induced, the observed parts of the data may
still hold much, if not all, of the information about the missing part
and the performance of the imputation procedure is not surprisingly
evaluated as very good. Also, it may be that the relations in real-life
data are rarely as pronounced as the problem that is being studied.

\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\hline
 & $\mu_{Y} = 10$ & $\beta_{X} = 0.4$ & CIW & CR & RMSE$_{p}$ & RMSE$_{c}$\\
\hline
\multicolumn{7}{l}{\textbf{\$n = 50\$}}\\
\hline
\hspace{1em}CCA & 9.961 & 0.341 & 0.823 & 0.94 & 0.866 & NaN\\
\hline
\hspace{1em}Imp. & 10.007 & 0.386 & 0.877 & 0.99 & 0.903 & 1.330\\
\hline
\hspace{1em}True & 10.021 & 0.393 & 0.541 & 0.97 & 0.903 & 0.000\\
\hline
\multicolumn{7}{l}{\textbf{\$n = 500\$}}\\
\hline
\hspace{1em}CCA & 9.941 & 0.343 & 0.232 & 0.85 & 0.897 & NaN\\
\hline
\hspace{1em}Imp. & 9.994 & 0.392 & 0.249 & 0.94 & 0.917 & 1.296\\
\hline
\hspace{1em}True & 9.998 & 0.400 & 0.161 & 0.90 & 0.914 & 0.000\\
\hline
\multicolumn{7}{l}{\textbf{\$n = 5000\$}}\\
\hline
\hspace{1em}CCA & 9.945 & 0.344 & 0.074 & 0.09 & 0.899 & NaN\\
\hline
\hspace{1em}Imp. & 10.000 & 0.397 & 0.078 & 0.91 & 0.916 & 1.298\\
\hline
\hspace{1em}True & 10.002 & 0.398 & 0.051 & 0.95 & 0.917 & 0.000\\
\hline
\end{tabu}

\hypertarget{missingness-generation-problems}{%
\subsection{Missingness generation
problems}\label{missingness-generation-problems}}

In order to obtain valid imputation inference, the imputation model must
capture the essence of the true non-response mechanism Meng (1994). The
model - if any - that is used to generate the missingness is usually
assumed to be random (MAR) or completely random (MCAR). Missingness
mechanisms that are not random (MNAR) and mechanisms that are considered
non-ignorable see e.g. Donald B. Rubin (1976), are generally ignored
during evaluation of imputation methodologies, except for those methods
that are specifically targeted at non-ignorable applications.

With MCAR missingness mechanisms, the probability to be missing is the
same for all cases. This is a necessary simulation condition for
evaluating the performance of imputation procedures. If an imputation
method is not able to solve the problem (i.e.~yield valid inference)
under MCAR, the statistical properties of the procedure are not sound.
Sadly, the straightforward case of MCAR is often neglected from
simulation studies and focus is drawn to the evaluation of MAR
mechanisms only. Alternatively, some studies limit their evaluations to
MCAR mechanisms only, which in my view may be far too simplistic.

\begin{figure}
\centering
\includegraphics{img/plot_mmech.pdf}
\caption{\label{fig:mis_mech}Three missingness mechanisms that yield
approximately 50 percent missingness (\(N=500\)). Displayed are a MCAR
mechanism, a right-tailed MAR mechanism generated from bivariate normal
data with correlation \(\rho=.8\) and the same MAR mechanism, now
generated from bivariate normal data with correlation \(\rho= 0\).}
\end{figure}

With MAR missingness mechanisms, the probability to be missing is the
same within groups of cases defined by the observed data only
(e.g.~males are less likely to disclose their weight, but gender is
observed). The essence here is that the observed relations in the data
are used to induce missingness during simulation (e.g.~weight is made
incomplete based on gender), but these relations may be weak, or
non-existent.

When MAR is induced based on weaker relations in the data, claims for a
method's applicability to situations where the missingness is random
become less valid. When MAR is induced from data without multivariate
relations, the inferential implication of the missingness would mimic
those of MCAR. This problem is generally overlooked in simulation
studies and the procedure to generate missing data is often
insufficiently described in resulting publications. Figure
\ref{fig:mis_mech} demonstrates two realizations of missingness
mechanisms under MAR mechanisms derived from data with different
multivariate relations.

It is obvious that the MAR missingness generated from data with low
correlation would yield similar inference to the MCAR missingness, even
though the missingness is clearly random. The simulation results from
Table 1 confirm this. A detailed simulation setup can be found in
Appendix A.

As expected, under genuine MAR missingness, i.e.~the missingness
probabilities differ conditional on the level of the incomplete
variable, the performance of complete case analyses drops dramatically.
For \emph{spurious} MAR, i.e.~the conditional probabilities remain the
same over the level of the incomplete variable, the performance is
comparable to that under MCAR. In fact, this would amount to one of the
special cases under which complete case analysis would be more efficient
than multiple imputation Van Buuren (2012), p.~48: the missingness does
not depend on the incomplete variable. This property might be useful in
practice, but considering it as a condition to evaluate performance
under MAR missingness is useless.

Although one cannot definitively verify if the missingness is random -
after all, for every MNAR mechanism there is a MAR mechanism with equal
fit Molenberghs et al. (2008) - it can be argued that MNAR is the more
likely mechanism for real life missingness scenarios. In such cases, an
indication of the validity of the obtained inference, given that the
assumed missingness mechanism is suspected to be invalid, may be
obtained by performing sensitivity analysis see e.g. Molenberghs et al.
(2014), part 5.

The amount of missingness that is used to simulate the performance of
imputation methods differs between studies. Some studies use only 10\%
missing data where other studies push the limits to additionally
investigate the performance under at least 50\% missing data. This
inconsistent display of simulation results may impact the objectivity of
meta-evaluations over imputation methods, as one method's performance
may appear to be favorable because of the less stringent simulation
conditions. This ultimately may lead to statisticians recommending a
less efficient method to applied researchers, thereby limiting the
efficiency of the imputation approach and unnecessarily lowering the
statistical power.

\hypertarget{performance-evaluation}{%
\subsection{Performance evaluation}\label{performance-evaluation}}

The evaluation criteria used to assess imputation performance vary from
one developer to another. This is not surprising as people from
different fields could have a different focus on the problem at hand.
However, because multiple imputation was designed as a mode for
statistical inference, a minimal set of statistical properties should at
least be evaluated. I highlight four aspects of evaluation that should
be considered when evaluation imputation routines.

First, developers often only inquire about the `accuracy' (i.e.~how well
can the method reproduce the original data). The goal of multiple
imputation is not to reproduce the data, but to allow for obtaining
valid inference given that the data are incomplete. We are interested in
the correct answer to the research question; not in the truth itself.
This means that, given the framework provided by Donald B. Rubin (1987),
statistical properties such as bias, confidence intervals and the
coverage rate of the confidence intervals should be studied. After all,
the 95\% confidence interval should contain the `true' value at least 95
out of 100 times Neyman (1934), p.~591.

Second, the performance of imputation procedures on distributional
properties is often ignored in simulation studies, and even though the
estimates on the analysis level may be justified, some methods can yield
imputations that may seem completely invalid to applied researchers. For
example, one could very accurately estimate average human height by
filling in negative values and values that are unrealistically large.
While the obtained inference could still be valid under such
imputations, the plausibility of the imputed values given the observed
data should be under scrutiny. Rather, one would prefer an imputation
technique to yield both valid inference and plausible imputations. It
should be studied if an imputation method is prone to deliver such
impractical results, and if so, under what conditions.

Third, sampling variation has always been an essential part of the
evaluation of multiple imputation methodology. However, in order to
obtain information about a method's ability to handle the missing data
problem, or to objectively compare methods on their ability to correct
for missingness, it is not necessary to take sampling variation into
account Vink and van Buuren (2014). After all, we are interested only in
the missing data mechanism, and are not considering the noise induced by
the sampling mechanism for evaluation in such studies.

Last, when the evaluator is on the verge of drawing conclusions about
the performance of the imputation routine, the performance should be
carefully qualified. Comparing the performance of an imputation routine
given a population (or true) parameter allows for quantitative
evaluation. However, in order to pose qualitative statements about the
performance on simulated conditions, comparative methodology is
required. For example, when claiming that imputation performance is
unacceptable when deviations from normality become rather stringent,
such performance is highly dependent on the simulation conditions that
are used. For a well-balanced judgement about the severity of the
performance drop, comparative simulations with e.g.~nonparametric models
should be executed. A method may perform badly, but if it still
outperforms every other approach, it may yet be of great practical
relevance.

\hypertarget{suggested-course-of-action}{%
\section{Suggested course of action}\label{suggested-course-of-action}}

\hypertarget{step-1.-obtain-truth}{%
\subsection{Step 1. Obtain truth}\label{step-1.-obtain-truth}}

The evaluator must to decide whether there is a need for sampling
variance in the simulations scheme. If sampling variance cannot be
omitted from the simulation scheme, two approaches are possible:

\begin{itemize}
\tightlist
\item
  Model-based simulation: draw samples from a known probability
  distribution, such as the multivariate normal distribution. The
  theoretical parameters under which the samples are obtained will serve
  as the comparative truth during the simulations.
\item
  Design-based simulation: sample data from a sufficiently large set.
  The parameters of the sufficiently large set will serve as the
  comparative truth during the simulations. This approach is often used
  in situations where a probability distribution is not available, or
  where real-life data structures are of interest. Applications of
  design-based simulation are often found in official statistics. A
  benefit of design-based simulation is the ability to use real-life
  observed data structures.
\end{itemize}

For situations where sampling variance is not of interest, the sampling
process can be omitted and a single complete dataset can simply be
obtained by model- or design-based approaches. The parameters of that
single complete set will serve as comparative truth. This process is
computationally convenient, because only a single complete datasets has
to be considered during all of the simulations. One needs to realize,
however, that the conventional pooling rules cf. Donald B. Rubin (1987),
pp.~76-77, do not apply for finite population inference and that
alternative pooling rules need to be used Raghunathan, Reiter, and Rubin
(2003), Vink and van Buuren (2014).

\hypertarget{step-2.-induce-missingness}{%
\subsection{Step 2. Induce
missingness}\label{step-2.-induce-missingness}}

First, one should always consider MCAR missingness, i.e.~the scenario
where the missing values are missing at random and the observed values
are observed at random. Under MCAR, the statistical properties of the
observed data given the missing data are known and any imputation
routine that cannot at least mimic the performance of the observed data
inference, should be deemed inefficient in the scope of the simulation.

Next, missing data should be induced conform a model that is dependent
on the observed data. A straightforward technique for inducing different
forms of univariate MAR missingness is described in Van Buuren (2012),
p.~63; see Figure \ref{fig:MAR} and a generalization to multivariate MAR
missingness can be found in Buuren et al. (2006), Appendix B, and Brand
(1999), par. 5.2.3. If the missingness is to be induced in longitudinal
data, auto-regressive MAR models e.g.~cf. Shara et al. (2015), model 2
and model 3, can be useful.

\begin{figure}
\centering
\includegraphics{img/plot_mar.pdf}
\caption{\label{fig:MAR}Examples of four random missingness mechanisms
that yield approximately 50 percent missingness. For negatively
correlated missingness covariates the resulting mechanisms are
reversed.}
\end{figure}

Third, it is advisable to investigate varying shapes of MAR missingness
to achieve a more realistic indication of the robustness of the
imputation performance across the range of random missingness. Given the
simulated data-distributions, one random missingness model may be far
more disastrous to the observed information than another model. This may
influence the performance of some (but not necessarily all) imputation
routines For example, inference from hot-deck techniques such as
predictive mean matching Roderick JA Little (1988), Donald B. Rubin
(1986) may be more severely impacted by large amounts of one-tailed
missingness than inference from parametric techniques. It would be a
shame to overlook such results due to the focus on a single MAR
mechanism.

Fourth, the amount of missingness must be varied. Remember that
missingness is only ignorable under MAR when the parameter of the data
is distinct and a-priori independent from the parameter of the missing
data process. Under MAR missingness we assume that we may use the
observed data to make inferences about the joint (observed and
unobserved) data.

The dependency of the procedure on the assumption under which we obtain
inference is only influenced by the amount of missingness. If there is
no missingness - or if there is no data, for that matter - the inference
does not depend on the assumption. Alternatively, the validity of
assumptions become increasingly important when the missingness
increases. Since we control the MAR mechanism, the assumption under
which we may solve the missing data problem should hold and it is only
fair to assess performance under stringent missingness conditions. I
therefor propose, for all mechanisms, to evaluate at least the following
univariate missingness percentages when evaluating imputation routines:

\begin{itemize}
\tightlist
\item
  10\% missingness: Depending on the size of the data, this percentage
  can be considered as a lower bound of realistic evaluation. Anything
  less than 10\% may be of little influence on the true data inference.
  Performance of a missing data method should at least be acceptable for
  most missing data problems.
\item
  25\% missingness: This is a fair amount of missingness and will,
  depending on the observed data information, have a noticeable
  influence on the completed data inference. When compared to the
  condition with 10\% missingness, the inference obtained under 25\%
  missingness should be less certain (i.e.~confidence/credibility
  interval width should increase), but estimates should still be
  properly covered and the statistical properties of the missing data
  method should be sound. In practice, at least to my experience in
  social sciences and official statistics, 25\% univariate missingness
  can easily be considered as a realistic missingness percentage.
\item
  50\% missingness: Performance under 50\% percent simulated missingness
  will most likely be impacted severely. Depending on how the
  missingness mechanism interacts with the simulated data, some
  imputation techniques may yield estimates that are under-covered such
  that the completed data inference should not be deemed valid anymore.
  If a method yields acceptable inference under 50\% MAR missingness, we
  can determine that the statistical properties of the imputation
  methodology are sound.
\end{itemize}

Although I limit the focus here to ignorable non-response, the above
suggested proportions are equally applicable to simulations under
non-ignorable non-response.

Fifth, the used missingness mechanism should be detailed, either
graphically or written as a function of the data. Often, when inducing
missingness, authors remain vague about the actual missingness mechanism
under investigation and, even worse, some authors only report something
like

\begin{quote}
\emph{We generated missing data following a MAR missingness mechanism.}
\end{quote}

This should be considered unacceptable as claims about the validity of
the multiple imputation inference depends heavily on the simulated
missingness mechanism.

\hypertarget{step-3.-evaluate-performance}{%
\subsection{Step 3. Evaluate
performance}\label{step-3.-evaluate-performance}}

It is wise to evaluate the performance of complete case analysis (aka
list-wise deletion) in all simulated conditions. We know the theoretical
properties of complete case analysis, which makes the technique useful
as a point of departure when evaluating multiple imputation performance.
The evaluation criteria may depend on the estimand. When descriptive
statistics are the goal and when statistical inference would not be of
interest, bias of the estimates would still apply, but standard errors
are generally ignored. Chances are that for those who focus on
descriptive statistical applications, multiple imputation would not be
the mode of choice. In general, I would say that each multiple
imputation routine should be evaluated on at least the following points:

\begin{itemize}
\tightlist
\item
  \emph{Bias:} Results should preferable be unbiased. However, the way
  bias is considered can greatly influence the interpretation of the
  results. For example, a negligible absolute bias for a parameter for
  which the true value is zero, would yield infinite bias when relative
  bias is considered. The way bias is computed should therefore be
  carefully chosen and described.
\item
  \emph{Interval width:} The way the confidence interval is calculated
  should be described. Wider intervals are associated with more
  uncertainty and the more narrow interval that is still properly
  covered indicates a sharper inference. However, inference from a wider
  interval that is properly covered is to be considered more valid when
  compared to a more narrow interval that is not properly covered
  anymore.
\item
  \emph{Coverage:} Coverage of a 95\% interval should in theory be
  \(\geq 95\), where a coverage of 95\% would be most efficient.
  Under-coverage (when estimation is too liberal) may be an indication
  of invalid inference, while over-coverage (when estimation is too
  conservative) tells us that efficiency could still be gained.
\item
  \emph{Distributional characteristics:} In practice, the distribution
  of the incomplete data may differ greatly from the observed data.
  Under anything but the MCAR assumption, this can be expected. When
  evaluating imputations, the distributional shapes should be checked
  and diagnostic evaluations should be performed see Abayomi, Gelman,
  and Levy (2008) for an detailed overview of diagnostic evaluation for
  multivariate imputations. When anomalies are found, and if the
  imputation method is valid, there should be an explanation, especially
  in the controlled environment of a properly executed simulation study.
\item
  \emph{Plausibility of the imputed values:} Plausible imputations -
  imputations that could be real values if they had been observed - are
  not a necessary condition for obtaining valid inference. However, in
  practice, especially when the imputer and the analyst are different
  persons, plausibility of imputations may be a desired property. When
  evaluating imputation routines, the evaluator should mention whether
  the routine is prone to deliver implausible values.
\item
  \emph{Convergence of the algorithm:} Most contemporary imputation
  techniques rely on iterative algorithms, such as the Gibbs sampler,
  where some algorithms are critically considered to be possibly
  incompatible Gibbs samplers, PIGS, Li, Yu, and Rubin (2012). The
  convergence of all iterative algorithms should always be considered
  and if non-convergence is suspected, the inference resulting from the
  imputations should not be considered.
\end{itemize}

For evaluations of model-based simulations, it could be convincing to
demonstrate a method's applicability to real-world missing data
problems. This can, for example, be achieved by obtaining and imputing a
fully observed set of variables, wherein the missingness is mimicked
from a similar, incomplete set. Alternatively, an incomplete set could
be obtained and truth could be established by removing the incomplete
cases from the data. However, the real-world missingness would then be
omitted.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

This document is aimed at establishing a common ground for the
evaluation of imputation routines. Such a common ground would be the
basis of a standardized evaluation. This allows for fair and efficient
comparisons between imputation techniques. Ultimately, it would be
desirable to evaluate every imputation routine against the same
standardized set in order to quantify the statistical properties across
imputation routines. If properly executed, this would allow for careful
matching of imputation methodologies to new missing data problems.

\hypertarget{appendix-a-simulation-setup}{%
\section{Appendix A: Simulation
setup}\label{appendix-a-simulation-setup}}

Let \(Y=(Y_{obs},Y_{mis})\) be an incomplete semi-continuous variable,
where \(Y_{obs}\) and \(Y_{mis}\) denote the observed values and the
missing values in \(Y\), respectively. Further, \(X=(X_1,...,X_k)\) is a
set of \(k\) fully observed covariates, where \(X_{obs}\) and
\(X_{mis}\) correspond to the observed an missing parts in \(Y\) and
where \(X_j\) would indicate the \(j\)th variable in \(X\), with
\(j=1,\dots, k\). Finally, let \(R\) be a response indicator that is 1
if \(Y\) is observed and 0 if \(Y\) is missing.

To prove the point, I generated a single data set from the multivariate
normal distribution with means \[
\mu= \bordermatrix{&     \cr
 Y  &5  \cr
{ X}_1  &10 \cr
{ X}_2  &10 },
\] and covariance matrix \[
\Sigma = \bordermatrix{& Y &{ X}_1 &{ X}_2   \cr
 Y  &1  &0.8    &0\cr
{ X}_1  &0.8    &1  &0\cr
{ X}_2  &0  &0  &1}.
\]

Three incomplete sets were then generated, wherein random missingness
was imposed in \(Y\) in three ways. First, I left the observed data
\emph{observed at random} to impose MCAR missingness following the
mechanism \[
P(R=0|Y_{obs},Y_{mis}, X_j)=P(R=0)=.50.
\] Next, I created random missingness in \(Y\) according to the
following MAR missingness mechanism \[
P(R=0|Y_{obs},Y_{mis}, X_j)=P(R=0|Y_{obs}, X_j),
\] by using a random draw from a binomial distribution of the same
length as \(Y\) and of size 1 with missingness probability equal to the
inverse logit \[
P(R=0)=\frac{e^{a}}{(1+e^{a})}.
\] where \(a=(\mu_{X_j}-X_{ij})/\sqrt{\mathbb{E}[X_{ij} - \mu_{X_j}]}\)
in order to create approximately 50\% right-tailed MAR missingness. I
followed this procedure twice; once for each \(X_j\).

The three resulting incomplete data sets were then imputed with package
\texttt{mice} Van Buuren and Groothuis-Oudshoorn (2011), v2.25, in
\texttt{R} Team (2015), v3.2.3, with Bayesian linear regression
imputation (\texttt{mice.impute.norm}) as the imputation routine. The
completed-data analyses were combined into a single inference following
the rules described in Vink and van Buuren (2014).

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{sessionInfo}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## R version 4.1.1 (2021-08-10)
## Platform: x86_64-w64-mingw32/x64 (64-bit)
## Running under: Windows 10 x64 (build 19042)
## 
## Matrix products: default
## 
## locale:
## [1] LC_COLLATE=Dutch_Netherlands.1252  LC_CTYPE=Dutch_Netherlands.1252   
## [3] LC_MONETARY=Dutch_Netherlands.1252 LC_NUMERIC=C                      
## [5] LC_TIME=Dutch_Netherlands.1252    
## 
## attached base packages:
## [1] stats     graphics  grDevices utils     datasets  methods   base     
## 
## other attached packages:
## [1] dplyr_1.0.7
## 
## loaded via a namespace (and not attached):
##  [1] Rcpp_1.0.7        highr_0.9         pillar_1.6.4      compiler_4.1.1   
##  [5] tools_4.1.1       digest_0.6.29     viridisLite_0.4.0 evaluate_0.14    
##  [9] lifecycle_1.0.1   tibble_3.1.6      lattice_0.20-44   pkgconfig_2.0.3  
## [13] rlang_0.4.12      rstudioapi_0.13   DBI_1.1.2         yaml_2.2.1       
## [17] mvtnorm_1.1-3     xfun_0.29         fastmap_1.1.0     kableExtra_1.3.4 
## [21] xml2_1.3.3        httr_1.4.2        withr_2.4.3       stringr_1.4.0    
## [25] knitr_1.37        systemfonts_1.0.3 generics_0.1.1    vctrs_0.3.8      
## [29] webshot_0.5.2     rprojroot_2.0.2   grid_4.1.1        tidyselect_1.1.1 
## [33] svglite_2.0.0     glue_1.6.0        mice_3.14.0       here_1.0.1       
## [37] R6_2.5.1          fansi_0.5.0       rmarkdown_2.11    purrr_0.3.4      
## [41] tidyr_1.1.4       magrittr_2.0.1    codetools_0.2-18  scales_1.1.1     
## [45] backports_1.4.1   ellipsis_0.3.2    htmltools_0.5.2   rvest_1.0.2      
## [49] assertthat_0.2.1  colorspace_2.0-2  utf8_1.2.2        stringi_1.7.6    
## [53] munsell_0.5.0     broom_0.7.12      crayon_1.4.2
\end{verbatim}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-abayomi2008diagnostics}{}}%
Abayomi, Kobi, Andrew Gelman, and Marc Levy. 2008. {``Diagnostics for
Multivariate Imputations.''} \emph{Journal of the Royal Statistical
Society: Series C (Applied Statistics)} 57 (3): 273--91.

\leavevmode\vadjust pre{\hypertarget{ref-alkire2015global}{}}%
Alkire, Blake C, Nakul P Raykar, Mark G Shrime, Thomas G Weiser, Stephen
W Bickler, John A Rose, Cameron T Nutt, et al. 2015. {``Global Access to
Surgical Care: A Modelling Study.''} \emph{The Lancet Global Health} 3
(6): e316--23.

\leavevmode\vadjust pre{\hypertarget{ref-allan2015determinants}{}}%
Allan, Stephen, and Julien Forder. 2015. {``The Determinants of Care
Home Closure.''} \emph{Health Economics} 24 (S1): 132--45.

\leavevmode\vadjust pre{\hypertarget{ref-brand1999development}{}}%
Brand, Jaap. 1999. {``Development, Implementation and Evaluation of
Multiple Imputation Strategies for the Statistical Analysis of
Incomplete Data Sets.''} PhD thesis, Erasmus University Rotterdam.

\leavevmode\vadjust pre{\hypertarget{ref-buur06}{}}%
Buuren, S. Van, J. P. L. Brand, C. G. M. Groothuis-Oudshoorn, and D. B.
Rubin. 2006. {``Fully Conditional Specification in Multivariate
Imputation.''} \emph{Journal of Statistical Computation and Simulation}
76 (12): 1049--64. \url{https://doi.org/10.1080/10629360600810434}.

\leavevmode\vadjust pre{\hypertarget{ref-crowley2014flexible}{}}%
Crowley, Jocelyn Elise, and Stanislav Kolenikov. 2014. {``Flexible Work
Options and Mothers' Perceptions of Career Harm.''} \emph{The
Sociological Quarterly} 55 (1): 168--95.

\leavevmode\vadjust pre{\hypertarget{ref-dafoe2013democratic}{}}%
Dafoe, Allan, John R Oneal, and Bruce Russett. 2013. {``The Democratic
Peace: {Weighing} the Evidence and Cautious Inference.''}
\emph{International Studies Quarterly} 57 (1): 201--14.

\leavevmode\vadjust pre{\hypertarget{ref-de2014religiosity}{}}%
De Hoon, Sean, and Frank Van Tubergen. 2014. {``The Religiosity of
Children of Immigrants and Natives in England, Germany, and the
Netherlands: {The} Role of Parents and Peers in Class.''} \emph{European
Sociological Review}, jcu038.

\leavevmode\vadjust pre{\hypertarget{ref-klausch2015selection}{}}%
Klausch, Thomas, Joop Hox, and Barry Schouten. 2015. {``Selection Error
in Single-and Mixed Mode Surveys of the {Dutch} General Population.''}
\emph{Journal of the Royal Statistical Society: Series A (Statistics in
Society)} 178 (4): 945--61.

\leavevmode\vadjust pre{\hypertarget{ref-li2012imputing}{}}%
Li, Fan, Yaming Yu, and Donald B Rubin. 2012. {``Imputing Missing Data
by Fully Conditional Models: {Some} Cautionary Examples and
Guidelines.''} \emph{Duke University Department of Statistical Science
Discussion Paper} 1124.

\leavevmode\vadjust pre{\hypertarget{ref-little2012calibrated}{}}%
Little, Roderick J. 2012. {``Calibrated {Bayes}, an Alternative
Inferential Paradigm for Official Statistics.''} \emph{Journal of
Official Statistics} 28 (3): 309.

\leavevmode\vadjust pre{\hypertarget{ref-little1988missing}{}}%
Little, Roderick JA. 1988. {``Missing-Data Adjustments in Large
Surveys.''} \emph{Journal of Business \& Economic Statistics} 6 (3):
287--96.

\leavevmode\vadjust pre{\hypertarget{ref-little2012prevention}{}}%
Little, Roderick J, Ralph D'Agostino, Michael L Cohen, Kay Dickersin,
Scott S Emerson, John T Farrar, Constantine Frangakis, et al. 2012.
{``The Prevention and Treatment of Missing Data in Clinical Trials.''}
\emph{New England Journal of Medicine} 367 (14): 1355--60.

\leavevmode\vadjust pre{\hypertarget{ref-meng94}{}}%
Meng, Xiao-Li. 1994. {``Multiple-{Imputation Inferences} with
{Uncongenial Sources} of {Input}.''} \emph{Statistical Science} 9 (4):
538--58. \url{https://doi.org/10.1214/ss/1177010269}.

\leavevmode\vadjust pre{\hypertarget{ref-molenberghs2008every}{}}%
Molenberghs, Geert, Caroline Beunckens, Cristina Sotto, and Michael G
Kenward. 2008. {``Every Missingness Not at Random Model Has a
Missingness at Random Counterpart with Equal Fit.''} \emph{Journal of
the Royal Statistical Society: Series B (Statistical Methodology)} 70
(2): 371--88.

\leavevmode\vadjust pre{\hypertarget{ref-molenberghs2014handbook}{}}%
Molenberghs, Geert, Garrett Fitzmaurice, Michael G Kenward, Anastasios
Tsiatis, and Geert Verbeke. 2014. \emph{Handbook of Missing Data
Methodology}. {CRC Press}.

\leavevmode\vadjust pre{\hypertarget{ref-morr18}{}}%
Morris, Tim P., Ian R. White, and Michael J. Crowther. 2019. {``Using
Simulation Studies to Evaluate Statistical Methods.''} \emph{Statistics
in Medicine} 38 (11): 2074--2102.
\url{https://doi.org/10.1002/sim.8086}.

\leavevmode\vadjust pre{\hypertarget{ref-neym34}{}}%
Neyman, Jerzy. 1934. {``On the {Two Different Aspects} of the
{Representative Method}: {The Method} of {Stratified Sampling} and the
{Method} of {Purposive Selection}.''} \emph{Journal of the Royal
Statistical Society} 97 (4): 558--625.
\url{https://doi.org/10.2307/2342192}.

\leavevmode\vadjust pre{\hypertarget{ref-pulkki2015cumulative}{}}%
Pulkki-Råback, Laura, Marko Elovainio, Christian Hakulinen, Jari
Lipsanen, Mirka Hintsanen, Markus Jokela, Laura D Kubzansky, et al.
2015. {``Cumulative Effect of Psychosocial Factors in Youth on Ideal
Cardiovascular Health in Adulthood the Cardiovascular Risk in Young
Finns Study.''} \emph{Circulation} 131 (3): 245--53.

\leavevmode\vadjust pre{\hypertarget{ref-raghunathan2003multiple}{}}%
Raghunathan, Trivellore E, Jerome P Reiter, and Donald B Rubin. 2003.
{``Multiple Imputation for Statistical Disclosure Limitation.''}
\emph{Journal of Official Statistics} 19 (1): 1.

\leavevmode\vadjust pre{\hypertarget{ref-rubin1986statistical}{}}%
Rubin, Donald B. 1986. {``Statistical Matching Using File Concatenation
with Adjusted Weights and Multiple Imputations.''} \emph{Journal of
Business \& Economic Statistics} 4 (1): 87--94.

\leavevmode\vadjust pre{\hypertarget{ref-rubi76}{}}%
Rubin, Donald B. 1976. {``Inference and {Missing Data}.''}
\emph{Biometrika} 63 (3): 581--92.
\url{https://doi.org/10.2307/2335739}.

\leavevmode\vadjust pre{\hypertarget{ref-rubi87}{}}%
---------. 1987. \emph{Multiple {Imputation} for Nonresponse in
Surveys}. Wiley Series in Probability and Mathematical Statistics
{Applied} Probability and Statistics. {New York, NY}: {Wiley}.

\leavevmode\vadjust pre{\hypertarget{ref-shara2015randomly}{}}%
Shara, Nawar, Sayf A Yassin, Eduardas Valaitis, Hong Wang, Barbara V
Howard, Wenyu Wang, Elisa T Lee, and Jason G Umans. 2015. {``Randomly
and Non-Randomly Missing Renal Function Data in the Strong Heart Study:
A Comparison of Imputation Methods.''} \emph{PloS One} 10 (9): e0138923.

\leavevmode\vadjust pre{\hypertarget{ref-sorbi2014medium}{}}%
Sorbi, MJ, AM Kleiboer, HG van Silfhout, G Vink, and J Passchier. 2014.
{``Medium-Term Effectiveness of Online Behavioral Training in Migraine
Self-Management: {A} Randomized Trial Controlled over 10 Months.''}
\emph{Cephalalgia : An International Journal of Headache},
0333102414547137.

\leavevmode\vadjust pre{\hypertarget{ref-stinesen2015reduced}{}}%
Stinesen Kollberg, Karin, Ann-Charlotte Waldenström, Karin Bergmark,
Gail Dunberger, Anna Rossander, Ulrica Wilderäng, Elisabeth
Åvall-Lundqvist, and Gunnar Steineck. 2015. {``Reduced Vaginal
Elasticity, Reduced Lubrication, and Deep and Superficial Dyspareunia in
Irradiated Gynecological Cancer Survivors.''} \emph{Acta Oncologica} 54
(5): 772--79.

\leavevmode\vadjust pre{\hypertarget{ref-R}{}}%
Team, R Core. 2015. \emph{R: {A} Language and Environment for
Statistical Computing}. Manual. {Vienna, Austria}: {R Foundation for
Statistical Computing}.

\leavevmode\vadjust pre{\hypertarget{ref-fimd}{}}%
Van Buuren, Stef. 2012. \emph{Flexible Imputation of Missing Data}. {CRC
press}.

\leavevmode\vadjust pre{\hypertarget{ref-mice}{}}%
Van Buuren, Stef, and Karin Groothuis-Oudshoorn. 2011. {``Mice:
{Multivariate Imputation} by {Chained Equations} in {R}.''}
\emph{Journal of Statistical Software} 45 (1): 1--67.
\url{https://doi.org/10.18637/jss.v045.i03}.

\leavevmode\vadjust pre{\hypertarget{ref-vink14}{}}%
Vink, Gerko, and Stef van Buuren. 2014. {``Pooling Multiple Imputations
When the Sample Happens to Be the Population,''} September.

\leavevmode\vadjust pre{\hypertarget{ref-wong2015economic}{}}%
Wong, Frances Kam Yuet, Ching So, June Chau, Antony Kwan Pui Law,
Stanley Ku Fu Tam, and Sarah McGhee. 2015. {``Economic Evaluation of the
Differential Benefits of Home Visits with Telephone Calls and Telephone
Calls Only in Transitional Discharge Support.''} \emph{Age and Ageing}
44 (1): 143--47.

\leavevmode\vadjust pre{\hypertarget{ref-zhao22}{}}%
Zhao, Yang. 2022. {``Diagnostic Checking of Multiple Imputation
Models.''} \emph{AStA Advances in Statistical Analysis}, January.
\url{https://doi.org/10.1007/s10182-021-00429-1}.

\end{CSLReferences}

\end{document}