### 2022-03-22 Simulation/Simulation.Rmd
Manuscript after discussion on suggestions Hanne, including first version of the checklist.

---
title: "Simulation study"
author: "Hanne Oberman"
date: "4-3-2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document contains the code and results for a simulation study accompanying the Standardized Evaluation manuscript.

# Overview

The simulation contains a whopping set of `r 3*3*3*4*3*2*4*3` methods/conditions. That's because we use different parameters for each level of the simulation, see pseudo-code:

```
for (data geneneration conditions) {
  generate complete data
  for (amputation conditions) {
    ampute complete data
    for (imputation conditions) {
      impute incomplete data
      analyze imputed data
      }}}
```

The conditions are as follows.

Data generation conditions:

- number of observations ($n = \{50, 500, 5000\}$) *[for efficiency: only max required?]*
- number of variables ($p = \{2\}$)
- coherence between variables ($r = \{0, 0.4, 0.8\}$)

Amputation conditions:

- missingness mechanism (`mech` = {MCAR, MAR, MNAR})
- type of M(N)AR (`type` = {LEFT, RIGHT, MID, TAIL})
- missingness proportion (`prop` = {0.1, 0.25, 0.5})

Imputation conditions:

- imputation method (`meth` = {none, mean imputation, regression imputation, stochastic regression})
- number of imputations (`m` = {1, 5}) *[for efficiency: only max required?]*
- number of iterations (`maxit` = 1, 5, 10) *[for efficiency: use mice.mids to shorten runtime?]*

Performance evaluation:
- estimand (average, regression coefficient)
- diagnostic (bias, coverage rate, confidence interval width, cell rmse, analysis model rmse)

# Simulation set-up

Set-up environment:

```{r env}
# random seed
set.seed(2022)
# packages
library(dplyr)
# utils functions
rmse <- function(obs, pred = 0){sqrt(mean((obs-pred)^2))}
```

## Simulation functions

For data generation we use the following:

```{r dat}
generation <- function(r = 0.4, n = 500, p = 2) {
  # generate data based on variance-covariance matrix
  vcov <- matrix(r, p, p)
  diag(vcov) <- 1
  dat <- data.frame(
    mvtnorm::rmvnorm(
    n,
    mean = rep(0, p),
    sigma = vcov
  )) %>% setNames(paste0("X", 1:p))
  # output
  out <- list(
    n_obs = n,
    n_var = p,
    corr = r,
    dat = dat)
  return(out)
}
```

For amputation of the complete data:

```{r amp}
amputation <- function(datlist, mech = "MAR", type = "MID", prop = 0.25) {
  # induce missingness
  amp <- mice::ampute(datlist$dat, 
                 prop = prop,
                 mech = mech,
                 type = type)
  # output
  out <- list(
    n_obs = datlist$n_obs,
    n_var = datlist$n_var,
    corr = datlist$corr,
    mis_mech = mech,
    mis_type = type,
    mis_prop = prop,
    amp = amp)
  return(out)
}
```

To impute the incomplete data:

```{r imp}
imputation <- function(amplist, meth = "norm", m = 5, it = 5) {
  # for CCA
  if(meth == "none") {
    imp <- na.omit(amplist$amp$amp)
    average <- mean(imp$X1)
    rmse_cell <- NA
  } 
  # for actual imputation methods
  if (meth != "none") {
    if (meth == "mean" | meth == "norm.predict") {
      imp <- mice::mice(
        amplist$amp$amp,
        m = 1,
        method = meth,
        maxit = 1,
        print = FALSE
      )
    } else {
    imp <- mice::mice(
        amplist$amp$amp,
        m = m,
        method = meth,
        maxit = it,
        print = FALSE
      )
    }
    average <- mice::complete(imp, "long") %>% .$X1 %>% mean()
    rmse_cell <- apply(imp$imp$X1 - amplist$amp$data$X1[imp$where[, "X1"]], 2, rmse) %>% mean()
  }
  # output
  out <- list(
    n_obs = amplist$n_obs,
    n_var = amplist$n_var,
    corr = amplist$corr,
    mis_mech = amplist$mis_mech,
    mis_type = amplist$mis_type,
    mis_prop = amplist$mis_prop,    
    imp_meth = meth,
    n_imp = m,
    n_it = it,
    average = average,
    rmse_cell = rmse_cell,
    imp = imp
  )
  return(out)
}
```

And finally, for performance evaluation:

```{r perf}
evaluation <- function(implist){
  # for CCA
  if (is.data.frame(implist$imp)) {
    fit <- lm(X1 ~ ., implist$imp)
    rmse_pred <- rmse(fit$residuals)
  }
  # for actual imputation methods
  if (mice::is.mids(implist$imp)) {
    mira <- mice::complete(implist$imp, "all") %>% 
      purrr::map(~lm(X1 ~ ., .))
    fit <- suppressWarnings(mice::pool(mira))
    rmse_pred <- purrr::map_dbl(mira, ~rmse(.$residuals)) %>% mean()
  }
  # get estimated regression coefficients
  est <- fit %>% broom::tidy(conf.int = TRUE) %>%
    .[2, c("estimate", "conf.low", "conf.high")] 
  # output
  out <- data.frame(
    n_obs = implist$n_obs,
    n_var = implist$n_var,
    corr = implist$corr,
    mis_mech = implist$mis_mech,
    mis_type = implist$mis_type,
    mis_prop = implist$mis_prop,    
    imp_meth = implist$imp_meth,
    n_imp = implist$n_imp,
    n_it = implist$n_it,
    average = implist$average, 
    coeff = est$estimate, 
    ci_ll = est$conf.low,
    ci_ul = est$conf.high,
    rmse_pred = rmse_pred, 
    rmse_cell = implist$rmse_cell
  )
  return(out)
}
```

The combined function will consist of these steps:

```{r test_default}
dat <- generation()
amp <- amputation(dat)
imp <- imputation(amp)
est <- evaluation(imp)
```

## Simulation conditions

The full simulation function will be applied to these conditions:

```{r cond}
# data generation parameters
n_obs = c(50, 500, 5000) 
corr = c(0, .4, .8)
dat_par <- expand.grid(n_obs = n_obs, corr = corr)
# amputation parameters
mis_mech = c("MCAR", "MAR", "MNAR")
mis_type = c("LEFT", "RIGHT", "MID", "TAIL")
mis_prop = c(.1, .25, .5)
amp_par <- expand.grid(dat_nr = 1:nrow(dat_par), mis_mech = mis_mech, mis_type = mis_type, mis_prop = mis_prop, stringsAsFactors = FALSE)
# imputation parameters
imp_meth = c("none", "mean", "norm.predict", "norm") 
n_imp = c(1, 5)
n_it = c(1, 5, 10)
imp_par <- expand.grid(amp_nr = 1:nrow(amp_par), imp_meth = imp_meth, n_imp = n_imp, n_it = n_it, stringsAsFactors = FALSE)
```

To generate the incomplete data for each condition we use:

```{r test_cond, eval=FALSE, include=FALSE}
all_dat <- purrr::pmap(dat_par, ~{generation(n = ..1, r = ..2)})
all_amp <- purrr::pmap(amp_par, ~{amputation(all_dat[[..1]], mech = ..2, type = ..3, prop = ..4)})
all_imp <- purrr::pmap_dfr(imp_par, ~{imputation(all_amp[[..1]], meth = ..2, m = ..3, it = ..4) %>% 
    evaluation()})
```