### 2022-11-23 Manuscript/response_reviewer_2.md

Reviewer 2

Thanks to the authors for addressing my comments on the first submission.

The authors choose to focus on the first option that I listed at the previous review: Incomplete covariate values in observational studies with focus on unbiased, efficient parameter estimation. This seems a sensible decision. This authors be made explicit in the title and abstract, both of which promise to be far more general.

- I noted a problem with a specific suggestion in section 2.1, which is to remove sampling variation by fixing the complete data and only simulating missing values. The authors have partially addressed this but I don???t think been clear enough. There are two important points that I think readers would miss:

  1. The suitability of this approach depends on the performance measures to be used. Specifically, any that involve sampling variation are of no interest (i.e. it???s only ok if you are only evaluating bias). The problem can be
2. This approach is restricted to using a selection-model factorisation when generating data. If using a pattern-mixture approach, you cannot do this. Perhaps this point is obvious but it???s not explicit.

**We have rewritten the paragraph to avoid confusion and have outlined that this scenario only applies to simulation studies where the simulator has the flexibility to induce missing data uncertainty in a finite population. In that case, the Rubin's rules about the total variance (Q-\bar{Q}) have to be adapted to T = B + B/m. We have also detailed this in the revised manuscript. If a finite population is repeatedly made incomplete in simulation, such a finite population may serve as comparative truth and valid inference are obtained by the above mentioned pooling rule. This scenario applies to all performance measures and data generating schemes and has been a part of our simulation toolbox for years. The publication by Vink & Van Buuren (2014) demonstrates application and validity.**

- In section 2.2, I like the recommendation to explore MNAR even if something is primarily intended to work with MAR. However, I think there may be two points to add. First, explicitly say that this is a way of understanding robustness: method A may perform ok under MAR and some MNAR mechanisms, while method B performs well under MAR but very poorly under any MNAR mechanism. In this case we may favour A over B. Second, this issue is IMO applicable to ???late-phase??? methodological research (see Heinze et al. https://arxiv.org/abs/2209.13358); in early-phase work, I think it's perfectly acceptable to say 'this method works well in the setting it's intended to work well in'.

**Thanks. This is a useful addition to this paragraph. We have taken and implemented your suggestions.**

- It took a while to get my head around figure 1, and I???d suggest a little more explanation. The ???MCAR??? and ???Right-tailed MAR (??=0.8)??? panels are clear, but the ???Right-tailed MAR (??=0)??? is not. First, I cannot see what is ???right-tailed??? about it. Second, while I agree that MCAR requires an identical probability of missingness for each individual, how is this realised in a simulation study involving repetitions? The right panel of figure 1 seems to be ???MAR given individual???. Are notional ???individuals??? retained for each repetition? I don???t know the answer here, but two points:
1. fixing something across repetitions vs. drawing it each time can make a difference (e.g. comment above about only simulating missing data).
2. If you choose not to fix this ???MAR??? mechanism across reps, it becomes operationally indistinguishable from MCAR.

**We understand the confusion as the right-tailed nature of the missingness is lost through the zero correlation between the variable to ampute and the variable that serves as missingness covariate. This is exactly the point of the figure, but we apparently have not made this sufficiently clear. Ww have adjusted the text and have added a table that includes a small set of evaluation metrics that demonstrate both the importance of valid missingness and the ability to infer invalid missingness mechanisms through the comparison between results obtained on MAR/MNAR with CCA and multiple imputation.**

- I like the properties suggested in section 2.4. Two comments:
1. It would be good to note that these properties are for an estimand of interest rather than just any parameter in the analysis model
2. (Root) mean squared error is arguably not for predictive accuracy, which you have said is not the target here, but a trade-off between bias and variance when a biased procedure has lower variance than an unbiased procedure.

**We've noted (1) and changed predictive accuracy to 'resemblence to the true values' for (2)**

  - In table 1, I wondered about adding a point under methods for ???how SEs and confidence intervals are to be constructed???. Rubin???s rules are likely the default but people might use Robins & Wang???s rules, Reiter???s rules, Efron???s full-mechanism bootstrap, von Hippel and Bartlett???s bootstrap???
**added**

  - The point about validity of imputations is interesting. Rubin makes the point that the aim of MI is validity of inference, not of imputations,  Do the authors have examples of how to assess the validity of imputations? I broadly agree with this as a useful diagnostic but ideas of how to operationalise it in simulation studies would be useful.
**We've added some suggestions to text and table**

MISC./MINOR COMMENTS
- The introduction says ???The idea behind imputation is to impute (fill in) missing values, to obtain a valid estimate of what could have been.??? This is behind some imputation but certainly not multiple imputation. Rubin explicitly talks about the idea as being valid inference, not to recreate the missing values.
**Although we fully agree with you, we believe this wording is correct as we do not yet speak of multiple imputation in particular and we talk about "valid estimate of what could have been" and not values that could have been. We also very clearly separate single and multiple imputation in the following paragraphs and do not wish to confuse the reader with semantics in the first paragraph.**

- The statement ???Real-life data hardly ever follow a known theoretical distribution??? seems impossible to falsify. Perhaps safer to describe this as ???We cannot know so safer not to pretend we do???.
**The revised manuscript relaxes this particular phrasing**

- Since reviewing the first submission, I've seen some simulation studies where people have MAR and positivity vs. MAR and non-positivity (or near non-positivity). No pressure but I think it might be worth mentioning this in section 2.3.
**It is not clear to us to what you refer with positivity. Do you refer by positivity to any unique combination of confounding variables having a non-zero probability of missingness? Or is your comment referring to (non)positive covariance matrices? That said, we realize that given the timeframe of the review process, many potentially interesting notes could be made about newly published or pre-printed research. References that redefined the state-of-the-art in imputation methodology evaluation or were vital to the validity of our manuscript, were added between review rounds. For reasons of brevity we cannot include all. If you feel that we have missed such a must-include reference, we would be delighted to hear from you. After all, your comments have been very valuable to us.**